Hierarchically modelling \emph{Kepler} dwarfs and subgiants to improve inference of stellar properties with asteroseismology

### 2021-02-02 paper/main.tex
HL: Fix typo
DH: Added references and acknowledgements

The inference of stellar ages, masses and radii has improved through the use of asteroseismology in recent years \citep[e.g. see the review by][]{Chaplin.Miglio2013}. Measuring the oscillation modes in stars using photometric time series data, from missions such as \emph{CoRoT} \citep{Baglin.Auvergne.ea2006}, \emph{Kepler} \citep{Borucki.Koch.ea2010} and \emph{TESS} \citep{Ricker.Winn.ea2015} has given us new insights into the structure and evolution of stars. Recent examples include a deeper understanding of stellar structure \citep{Verma.Raodeo.ea2017}, chronology of a Milky Way merger \citep{Chaplin.Serenelli.ea2020} and classifying exoplanetary systems \citep{Huber.Chaplin.ea2019}. Several studies used grids of stellar models with constraints from asteroseismology to produce catalogues of precise stellar parameters \citep{Pinsonneault.Elsworth.ea2014, SilvaAguirre.Lund.ea2017}. However, with increasing precision on fundamental parameters inferred from stellar models with asteroseismology, extra care should be taken to ensure that we are accounting for uncertainty in our choice of stellar physics.

@@

Our HBM requires a way to map from the stellar initial (or bulk) properties to predict observables. We can achieve this with a large grid of stellar evolutionary models. However, a discrete grid can produce inaccurate posterior distributions, limited to the grid resolution. Increasing the resolution is computationally demanding, especially when scaling to higher input dimensions. One solution is to interpolate the stellar models, as is common in the isochrone-fitting method \citep[see e.g.][]{Berger.Huber.ea2020}. However, interpolation can become computationally expensive at high input dimensions and grid size, and evaluating the likelihood using modern Bayesian sampling techniques is slow. Therefore, we use machine learning to map stellar model inputs to observables to provide a fast way to sample the HBM. In this work, we train an artificial neural network (ANN) on a large grid of stellar models. There have been similar applications of ANNs in asteroseismology \citep{Verma.Hanasoge.ea2016, Hendriks.Aerts2019} but not yet in the context of an HBM. Using the machine learning speed-up, we demonstrate a scalable method for obtaining fundamental stellar parameters.

@@

% Return here to see if these two paragraphs may be merged.
We determined luminosities, $L$ for the sample using the direct method of \textsc{isoclassify} \citep{Huber.Zinn.ea2017, Berger.Huber.ea2020} with \emph{K\textsubscript{S}}-band photometry, \emph{Gaia} DR2 parallaxes, ASPCAP $\metallicity$ and $\teff$ and asteroseismic $\log g$ as inputs. This involved computing absolute \emph{K\textsubscript{S}}-band magnitudes using the \emph{Gaia} DR2 parallaxes and extinctions determined by the 3D galactic reddening maps of \citet{Green.Schlafly.ea2018}. We determined absolute bolometric magnitudes by interpolating the MIST bolometric correction tables as a function of $\teff$, $\log g$ and $\metallicity$ \citep{Dotter2016, Choi.Dotter.ea2016}. We adopted the uncertainty of \SI{0.02}{\magnitude} assumed by \textsc{isoclassify} for both the extinctions and bolometric corrections, representative of typical systematics during interpolation \citep{Huber.Zinn.ea2017}. We obtained luminosities for the sample with a median uncertainty of 3.4 per cent.

@@

We built the train and validation dataset from the outputs of the grid of stellar models in Section \ref{sec:grid}. This included the input parameters: $M$, $\mlt$, $Y_\mathrm{init}$ and the initial heavy-elements fraction, $Z_\mathrm{init}$. We also included the $\teff$, $\log g$, $\dnu$, stellar age ($\tau$), radius ($R$), surface metallicity ($\metallicity_\mathrm{surf}$) and other chemical composition information generated by the models. We determined the fractional main sequence (MS) lifetime, $f_{\mathrm{MS}} = \tau / \tau_{\mathrm{MS}}$, of each evolutionary track by taking $\tau_{\mathrm{MS}}$ as the age when the central hydrogen fraction, $X_c < 0.01$. We then cut data where $f_{\mathrm{MS}} < 0.01$ to remove points on the grid prior to the MS. Once we had refined the data from the grid of models, we randomly sampled \num{7.736e6} points to use as the training dataset, with the remaining $\sim \num{2e6}$ points given to the validation dataset. We varied our choice of ANN input and output parameters among those available in the dataset during tuning (see Section \ref{sec:opt}).

@@

There are a few prior studies which look at the spread in $\mlt$ for a population of stars, typically by fitting $\mlt$ as a function of $\metallicity$, $\teff$ and $\log g$. For example, results from \citet{Viani.Basu.ea2018} for stellar models including diffusion, predict $\mlt$ in the range \numrange{1.5}{2.3} across our sample. This dispersion would be more compatible with the larger spread obtained by our PPS model. However, in future work we should further investigate how $\mlt$ varies with stellar parameters, as our assumption of a normal distribution may not be accurate.

@@

KIC 9025370 turned out to be a double-lined spectroscopic binary \citep{Nissen.SilvaAguirre.ea2017}, discovered after \citetalias{Serenelli.Johnson.ea2017} and hence included in the original sample. The brighter observed luminosity and possibly unreliable spectroscopic $\teff$ compared to our models were consistent with a spectroscopic binary. We calculated a photometric $\teff$ using the IRFM method \citep{Casagrande.Ramirez.ea2010} with the available 2MASS photometry for the target and obtained $\teff=\SI{5983(120)}{\kelvin}$, more consistent with our modelled effective temperature and inconsistent with its spectroscopic $\teff$. Thus, our inferred $\teff$ was within the dispersion between different observed $\teff$ scales. Running the model without KIC 9025370 did not affect the resulting inferred hyperparameters, demonstrating the robustness of our model. Therefore, we present KIC 9025370 in our results but suggest that further investigation should be carried out.

@@

This work is a part of a project that has received funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (CartographY; grant agreement ID 804752). We thank Ilya Mandel and Sean Matt for their discussion regarding this work.

>>>>

The inference of stellar ages, masses and radii has improved through the use of asteroseismology in recent years \citep[e.g. see the review by][]{Chaplin.Miglio2013}. Measuring the oscillation modes in stars using photometric time series data, from missions such as \emph{CoRoT} \citep{Baglin.Auvergne.ea2006}, \emph{Kepler} \citep{Borucki.Koch.ea2010} and \emph{TESS} \citep{Ricker.Winn.ea2015} has given us new insights into the structure and evolution of stars. Recent examples include a deeper understanding of stellar structure \citep{Verma.Raodeo.ea2017}, chronology of a Milky Way merger \citep{Chaplin.Serenelli.ea2020} and classifying exoplanetary systems \citep{Huber.Chaplin.ea2019, Tayar.Claytor.ea2020}. Several studies used grids of stellar models with constraints from asteroseismology to produce catalogues of precise stellar parameters \citep{Pinsonneault.Elsworth.ea2014, SilvaAguirre.Lund.ea2017}. However, with increasing precision on fundamental parameters inferred from stellar models with asteroseismology, extra care should be taken to ensure that we are accounting for uncertainty in our choice of stellar physics.

@@

Our HBM requires a way to map from the stellar initial (or bulk) properties to predict observables. We can achieve this with a large grid of stellar evolutionary models. However, a discrete grid can produce inaccurate posterior distributions, limited to the grid resolution. Increasing the resolution is computationally demanding, especially when scaling to higher input dimensions. One solution is to interpolate the stellar models, as is common in the isochrone-fitting method \citep[see e.g.][]{Berger.Huber.ea2020}. However, interpolation can become computationally expensive at high input dimensions and grid size, and evaluating the likelihood using modern Bayesian sampling techniques is slow. Therefore, we use machine learning to map stellar model inputs to observables to provide a fast way to sample the HBM. In this work, we train an artificial neural network (ANN) on a large grid of stellar models. There have been similar applications of ANNs in asteroseismology \citep{Verma.Hanasoge.ea2016, Bellinger.Angelou.ea2016, Hon.Stello.ea2017, Hon.Stello.ea2018a, Hendriks.Aerts2019} but not yet in the context of an HBM. Using the machine learning speed-up, we demonstrate a scalable method for obtaining fundamental stellar parameters.

@@

% Return here to see if these two paragraphs may be merged.
We determined luminosities, $L$ for the sample using the direct method of \textsc{isoclassify} \citep{Huber.Zinn.ea2017, Berger.Huber.ea2020} with \emph{K\textsubscript{S}}-band photometry, \emph{Gaia} DR2 parallaxes, ASPCAP $\metallicity$ and $\teff$ and asteroseismic $\log g$ as inputs. This involved computing absolute \emph{K\textsubscript{S}}-band magnitudes using the \emph{Gaia} DR2 parallaxes and extinctions determined by the 3D galactic reddening maps of \citet{Green.Schlafly.ea2018}. We determined absolute bolometric magnitudes by interpolating the MIST bolometric correction tables as a function of $\teff$, $\log g$ and $\metallicity$ \citep{Dotter2016, Choi.Dotter.ea2016}. We adopted the uncertainty of \SI{0.02}{\magnitude} assumed by \textsc{isoclassify} for both the extinctions and bolometric corrections, corresponding to typical systematic errors in extinction maps and bolometric fluxes \citep[e.g.][]{Zinn.Pinsonneault.ea2019a,Tayar.Claytor.ea2020}. We obtained luminosities for the sample with a median uncertainty of 3.4 per cent.

@@

We built the train and validation datasets from the outputs of the grid of stellar models in Section \ref{sec:grid}. This included the input parameters: $M$, $\mlt$, $Y_\mathrm{init}$ and the initial heavy-elements fraction, $Z_\mathrm{init}$. We also included the $\teff$, $\log g$, $\dnu$, stellar age ($\tau$), radius ($R$), surface metallicity ($\metallicity_\mathrm{surf}$) and other chemical composition information generated by the models. We determined the fractional main sequence (MS) lifetime, $f_{\mathrm{MS}} = \tau / \tau_{\mathrm{MS}}$, of each evolutionary track by taking $\tau_{\mathrm{MS}}$ as the age when the central hydrogen fraction, $X_c < 0.01$. We then cut data where $f_{\mathrm{MS}} < 0.01$ to remove points on the grid prior to the MS. Once we had refined the data from the grid of models, we randomly sampled \num{7.736e6} points to use as the training dataset, with the remaining $\sim \num{2e6}$ points given to the validation dataset. We varied our choice of ANN input and output parameters among those available in the dataset during tuning (see Section \ref{sec:opt}).

@@

There are a few prior studies which look at the spread in $\mlt$ for a population of stars, typically by fitting $\mlt$ as a function of $\metallicity$, $\teff$ and $\log g$ \citep[e.g.][]{Bonaca.Tanner.ea2012,Viani.Basu.ea2018}. For example, results from \citet{Viani.Basu.ea2018} for stellar models including diffusion, predict $\mlt$ in the range \numrange{1.5}{2.3} across our sample. This dispersion would be more compatible with the larger spread obtained by our PPS model. However, in future work we should further investigate how $\mlt$ varies with stellar parameters, as our assumption of a normal distribution may not be accurate.

@@

KIC 9025370 turned out to be a double-lined spectroscopic binary \citep{Nissen.SilvaAguirre.ea2017}, discovered after \citetalias{Serenelli.Johnson.ea2017} and hence included in the original sample. The higher observed luminosity from \textsc{Isoclassify} and inconsistent spectroscopic $\teff$ compared with our model posteriors were compatible with a spectroscopic binary. We calculated a photometric $\teff$ using the IRFM method \citep{Casagrande.Ramirez.ea2010} with the available 2MASS photometry for the target and obtained $\teff=\SI{5983(120)}{\kelvin}$, more consistent with our modelled effective temperature and inconsistent with its spectroscopic $\teff$. Thus, our inferred $\teff$ was within the dispersion between different observed $\teff$ scales. Running the model without KIC 9025370 did not affect the resulting inferred hyperparameters, demonstrating the robustness of our model. Therefore, we present KIC 9025370 in our results but suggest that further investigation should be carried out.

@@

This work is a part of a project that has received funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (CartographY; grant agreement ID 804752). We thank Ilya Mandel and Sean Matt for their discussion regarding this work. D.H. acknowledges support from the Alfred P. Sloan Foundation, the National Aeronautics and Space Administration (80NSSC19K0597), and the National Science Foundation (AST-1717000).
